One exact and one random sampling method to assess the reliability of inferences from network enrichment analysis

### 2018-10-25 doc/genesetdp.tex
items 5 and 6 of checklist
commented out figures

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{A simple null model for inferences from network enrichment analysis} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Gustavo S. Jeuken\textsuperscript{1} and
Lukas K\"{a}ll\textsuperscript{1,*}
\\
\bigskip
\textbf{1} Science for Life Laboratory, School of Engineering Sciences in Chemistry, Biotechnology and Health, KTH -- Royal Institute of Technology, Box 1031, 17121 Solna, Sweden
\\
\bigskip

@@

\begin{figure}[htb]
	\begin{center}
		\includegraphics[width=.9\textwidth]{figures/score_distribuition_multiple.png}
    \end{center}
  \caption{{\bf Score distribution of a query size of 10, 20 and 30 genes against the ``Glycolysis/Gluconeogenesis'' pathway.} GeneSetDP enables us to calculate the full score distribution for a query of a given size, $Q$, i.e. how many ways one can reach a certain score when adding up the scores for $Q$ genes. A red dot marks the maximum score, $S$, for each case.}
  \label{fig:score_dist}
\end{figure}

@@

\begin{figure}[htb]
  \begin{center}
		  \begin{tabular}[t]{c}
				\includegraphics[width=.9\textwidth]{figures/calibration_multiple.png} \\
				\includegraphics[width=0.45\textwidth]{figures/calibration_legend.png}
		\end{tabular}
  \end{center}
  \caption{{\bf Calibration of GeneSetDP and GeneSetMC.} We selected 10000 random sets of 10, 20 and 30 genes from the investigated genome definition and calculated their $p$ values with GeneSetMC and GeneSetDP against the ``Glycolysis/Gluconeogenesis'' pathway. To test the $p$~values uniformity we plotted them against their {\em normalized rank}, i.e. each $p$~value's (<rank>-0.5)/<total number of $p$~values>. We also added the calibration curve of BinoX, evaluated on the same random sets. The dashed line shows $y = x$ and the dotted lines $y = 2x$ and $y = 0.5x$, for comparison.}
  \label{fig:calibration}
\end{figure}

To test the consistency of the methods, we also investigated the correspondence between the $p$~values reported by GeneSetDP and GeneSetMC. As can be seen in Figure~\ref{fig:pscatter}, the results are consistent as expected.

\begin{figure}[htb]
  \begin{center}
		  \begin{tabular}[t]{c}
        % (Idea from reviewer comment in email 2018-06-08.)
				\includegraphics[width=.9\textwidth]{figures/scatter_dp_mc.png}
		\end{tabular}
  \end{center}
  \caption{{\bf Correspondence between GeneSetDP's and GeneSetMC's $p$~values.} We selected 10000 random sets of 10, 20 and 30 genes from the investigated genome definition and calculated their $p$ values with GeneSetMC and GeneSetDP against the ``Glycolysis/Gluconeogenesis'' pathway. We subsequently plotted the obtained $p$~values against each other. Again, the dashed line shows $y = x$ and the dotted lines $y = 2x$ and $y = 0.5x$.}
  \label{fig:pscatter}
\end{figure}

>>>>

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{A simple null model for inferences from network enrichment analysis} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Gustavo Jeuken\textsuperscript{1} and
Lukas K\"{a}ll\textsuperscript{1,*}
\\
\bigskip
\textbf{1} Science for Life Laboratory, School of Engineering Sciences in Chemistry, Biotechnology and Health, KTH -- Royal Institute of Technology, Box 1031, 17121 Solna, Sweden
\\
\bigskip

@@

\begin{figure}[htb]
	\begin{center}
		% \includegraphics[width=.9\textwidth]{figures/score_distribuition_multiple.png}
    \end{center}
  \caption{{\bf Score distribution of a query size of 10, 20 and 30 genes against the ``Glycolysis/Gluconeogenesis'' pathway.} GeneSetDP enables us to calculate the full score distribution for a query of a given size, $Q$, i.e. how many ways one can reach a certain score when adding up the scores for $Q$ genes. A red dot marks the maximum score, $S$, for each case.}
  \label{fig:score_dist}
\end{figure}

@@

\begin{figure}[htb]
  \begin{center}
		  \begin{tabular}[t]{c}
				% \includegraphics[width=.9\textwidth]{figures/calibration.png}
		\end{tabular}
  \end{center}
  \caption{{\bf Calibration of GeneSetDP and GeneSetMC.} We selected 10000 random sets of 10, 20 and 30 genes from the investigated genome definition and calculated their $p$ values with GeneSetMC and GeneSetDP against the ``Glycolysis/Gluconeogenesis'' pathway. To test the $p$~values uniformity we plotted them against their {\em normalized rank}, i.e. each $p$~value's (<rank>-0.5)/<total number of $p$~values>. We also added the calibration curve of BinoX, evaluated on the same random sets. The dashed line shows $y = x$ and the dotted lines $y = 2x$ and $y = 0.5x$, for comparison.}
  \label{fig:calibration}
\end{figure}

To test the consistency of the methods, we also investigated the correspondence between the $p$~values reported by GeneSetDP and GeneSetMC. As can be seen in Fig~\ref{fig:pscatter}, the results are consistent as expected.

\begin{figure}[htb]
  \begin{center}
		  \begin{tabular}[t]{c}
        % (Idea from reviewer comment in email 2018-06-08.)
				% \includegraphics[width=.9\textwidth]{figures/scatter_dp_mc.png}
		\end{tabular}
  \end{center}
  \caption{{\bf Correspondence between GeneSetDP's and GeneSetMC's $p$~values.} We selected 10000 random sets of 10, 20 and 30 genes from the investigated genome definition and calculated their $p$ values with GeneSetMC and GeneSetDP against the ``Glycolysis/Gluconeogenesis'' pathway. We subsequently plotted the obtained $p$~values against each other. Again, the dashed line shows $y = x$ and the dotted lines $y = 2x$ and $y = 0.5x$.}
  \label{fig:pscatter}
\end{figure}
